---
title: "EDA with commentary"
author: "Group E"
date: "4/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Load libraries

```{r load_libraries, message=FALSE}
# General purpose
library(tidyverse)
library(data.table)
library(lubridate)
library(DataExplorer) #gva

# Descriptive
library(skimr)

# Visualization
library(ggplot2)
library(PerformanceAnalytics)
library(corrplot)

# Mapping
library(leaflet)
library(leaflet.extras)
library(sf)
#library(gdtools)

# Calculations
library(forecast)
library(caret)
library(mice)
library(outliers) 

library(foreach)
library(doParallel)
```

## Source From R script
```{r}

source('20200411_Main_EDA.R', echo = FALSE, print.eval = FALSE)

```

## Read prepared Datasets
```{r eval=FALSE}
# Solar production dataset 
data_solar_train
data_solar_test
data_station 
data_add

```

# Overview: Data structure and content

## General overview

### NA check

```{r}
f_check_na(data_solar_train)
f_check_na(data_solar_test)
f_check_na(data_station)
f_check_na(data_add)

```
> Comment: The missing values in data_add will be imputed later

### Constant Variables check
```{r}
criteria_variables(data_solar_train)
criteria_variables(data_solar_test)
criteria_variables(data_station)
criteria_variables(data_add)
```

### Outliers check

Based on:
https://www.kaggle.com/rtatman/data-cleaning-challenge-outliers
https://cran.r-project.org/web/packages/outliers/outliers.pdf

#### Outliers in Solar data

```{r}

solar_data_outlier_skim
solar_data_outlier_hist
solar_data_outlier_table

# summary(unlist(data_solar_produ_scores))
# hist(unlist(data_solar_produ_scores))
# 
# table(
#   abs(unlist(data_solar_produ_scores))>=3
# )
```
> The data is normally distributed

#### Additional data
First, need to complete the data set to be able to use 'outl
iers()'.



