---
title: "EDA with commentary"
author: "Group E"
date: "4/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Load libraries

```{r load_libraries, message=FALSE}
# General purpose
library(tidyverse)
library(data.table)
library(lubridate)

# Descriptive
library(skimr)

# Visualization
library(ggplot2)
library(PerformanceAnalytics)
library(corrplot)

# Mapping
library(leaflet)
library(leaflet.extras)
library(sf)
#library(gdtools)

# Calculations
library(forecast)
library(caret)
library(mice)
library(outliers) 

library(foreach)
library(doParallel)
```

## Source From R script
```{r}

source('20200411_Main_EDA.R', echo = FALSE, print.eval = FALSE)

```

## Read prepared Datasets
```{r,eval=FALSE}
# Solar production dataset 
data_solar_train
data_solar_test
data_station 
data_add

```

# Overview: Data structure and content

## General overview

### NA check

```{r}
f_check_na(data_solar_train)
f_check_na(data_solar_test)
f_check_na(data_station)
f_check_na(data_add)

```
> Comment: The missing values in data_add will be imputed later

### Constant Variables check
```{r}
criteria_variables(data_solar_train)
criteria_variables(data_solar_test)
criteria_variables(data_station)
criteria_variables(data_add)
```

### Outliers check

Based on:
https://www.kaggle.com/rtatman/data-cleaning-challenge-outliers
https://cran.r-project.org/web/packages/outliers/outliers.pdf

#### Outliers in Solar data

```{r}

solar_data_outlier_skim
solar_data_outlier_hist
solar_data_outlier_table

```
> The data is normally distributed. There is only one row that could be considered outliers. 

#### Additional data
First, need to complete the data set to be able to use 'outliers()'.

## Solar dataset

### Training

```{r, eval=FALSE}
skim(data_solar_train)
glimpse(data_solar_train)
```

### Testing

```{r, eval=FALSE}
skim(data_solar_test)
glimpse(data_solar_test)
```

## Stations informaiton

```{r, eval=FALSE}
skim(data_station)
glimpse(data_station)
```

## Additional information

```{r, eval=FALSE}
skim(data_add)
glimpse(data_add)
```

# Descriptive plots

## Principal weather stations
First, rank weather stations sorted by volume
```{r}

principal_weather_station

```
Then, we create 2 plots, the first one shows top 7 weather stations
```{r}
principal_station_plot
```
## Rank position change over time
We are interested to find out how does the production changes over time. By looking at the top 10 stations. 

```{r}
p_rank
```
## Total production
Descriptive plot to show the production recorded in millions
```{r}
total_production_p
```

## Evolution of principal weather stations
Plot for the top highest weather stations.
```{r}
p_top_five_ts
```
## What about Training and Test?
```{r}
p_train_test
```

## Seasonality
https://towardsdatascience.com/forecasting-with-r-trends-and-seasonality-def24280e71f
```{r}
p_seasonality_all
```

### Seanonality of the top 5 stations

```{r}
p_seasonality_top_5
```

# Geolocalization

## Positions - Where are the stations located?
```{r}
map_stations
```

## Heatmap - visualizes production volume
```{r}
map_production
```

# Correlations 
## What can we see from the top 5 stations with the top 5 principal components?
```{r}
p_corr
```

# Histograms, density kernel and boxplot

## Of general production
```{r}
p_distri_prod
```

## Of each weather station 
```{r}
p_histogram_density

p_boxplot
```
## Of predictors 
```{r}
predictor_density
predictor_boxplot
```

# Variable Importance 

Computed Variable Importance is stored in data_solar_importance
```{r}
names(data_solar_importance)
```
> Note: used parallel computing to reduce run time 

# Additional information

## Feature visualization
```{r}
p_additional
```
## Correlation in additional dataset
```{r}
p_corrplot
```
## Impute missing values in additional dataset
```{r}
complete_mice
```
## Outliers in additional dataset
```{r}
add_outlier
add_outlier_hist
add_outlier_table 
```

## PCA of additional dataset

```{r}
data_add_col_pca
data_add_pca
```
### Plot density and boxplots
```{r}
pca_histogram_density
pca_boxplot

```
