---
title: "Programming R Workgroup Project: Machine Learning Model"
author: "Group E"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Load libraries

```{r load_libraries, message=FALSE}
# General porpuse
library(tidyverse)
library(data.table)
library(lubridate)
library(dplyr)

# Descriptive
library(skimr)

# Visualization
library(ggplot2)
library(ggpubr)

# Clustering
library(factoextra)
library(NbClust)

# Machine learning
library(e1071)
library(caret)

# Calculations
library(mice)

# Paralel computing
library(foreach)
library(doParallel)
```

# Definitions

```{r}
num_clusters <- 8
num_lag <- 30
num_times <- 5
num_col_importance <- 30
```

# Load data

```{r load_data}
data_solar <- readRDS(file = file.path('data', 'solar_dataset.RData'))
data_station <- fread(file = file.path('data', 'station_info.csv'))
data_add <- readRDS(file = file.path('data', 'additional_variables.RData'))
```

## Transform data

```{r}
# Source dataset
data_solar <- data_solar[j = Date2 := as.Date(x = Date, format = "%Y%m%d")]

# Add date conversions
data_solar <- data_solar %>% 
  mutate(Year = year(Date2),
         Month = lubridate::month(Date2, label = TRUE),
         Day = lubridate::day(Date2),
         Day_Of_Year = lubridate::yday(Date2),
         Day_Of_Week = lubridate::wday(Date2, label = TRUE, week_start = 1),
         Days_Since_Origin = time_length(interval(origin, Date2), unit = 'day')) %>% 
  as.data.table(.)

# Columns defined from the enunciate
data_solar_col_produ <- colnames(data_solar)[2:99]
data_solar_col_predi <- colnames(data_solar)[100:456]
data_solar_col_dates <- setdiff(colnames(data_solar), c(data_solar_col_produ, data_solar_col_predi))

# Columns defined from the enunciate
data_add_col <- colnames(data_add)[2:101]
data_add_col_dates <- setdiff(colnames(data_add), data_add_col)
```

## Complete data_add

```{r}
data <- select(data_add, all_of(data_add_col))

m_ <- 5
maxit_ <- 5
# data_mice_ <- mice(data, m=m_, maxit=maxit_, meth='pmm', seed=500)
# saveRDS(data_mice_, file.path('storage', 'data_add_mice.rds'))
data_mice_ <- readRDS(file.path('storage', 'data_add_mice.rds'))
# summary(data_mice_)

# Average of all the Multivariate Imputation
data_mice <- 0
for (i in 1:m_) data_mice <- data_mice + complete(data_mice_, i)
data_mice <- data_mice/m_

data_add_mice <- bind_cols(select(data_add, all_of(data_add_col_dates)), data_mice)

# Cleanup
rm(list = c('data', 'data_mice_', 'm_', 'maxit_', 'i', 'data_mice', 'data_add'))
```

# Join datasets

```{r}
data_solar_add <- data_solar %>% 
  left_join(data_add_mice, by = 'Date', suffix = c(".solar", ".add"))

rm(list = c('data_solar', 'data_add_mice'))
# skim(data_solar_add)
```

# Clustering
## Preparing data

```{r}
data_solar_add_train_ <- data_solar_add[1:5113, ]

# Merge data
data_solar_description <- data_solar_add_train_ %>% 
  select(all_of(data_solar_col_produ)) %>% 
  pivot_longer(cols = all_of(data_solar_col_produ), names_to = 'WeatherStation', values_to = 'Value') %>%
  group_by(WeatherStation) %>%
  summarize(mean = mean(Value), 
            sd = sd(Value), 
            q25 = quantile(Value, probs = .25),
            q50 = quantile(Value, probs = .5),
            q75 = quantile(Value, probs = .75),
            max = max(Value),
            min = min(Value),
            first = first(Value),
            last = last(Value))

data_solar_description_last_year <- data_solar_add_train_ %>% 
  filter(year(Date2) == year(max(Date2))) %>% 
  select(all_of(data_solar_col_produ)) %>% 
  pivot_longer(cols = all_of(data_solar_col_produ), names_to = 'WeatherStation', values_to = 'Value') %>%
  group_by(WeatherStation) %>%
  summarize(mean_last_year = mean(Value), 
            sd_last_year = sd(Value), 
            q25_last_year = quantile(Value, probs = .25),
            q50_last_year = quantile(Value, probs = .5),
            q75_last_year = quantile(Value, probs = .75),
            max_last_year = max(Value),
            min_last_year = min(Value),
            first_last_year = first(Value))

data_solar_description <- data_solar_description %>% 
  inner_join(data_solar_description_last_year, by = 'WeatherStation') %>% 
  inner_join(data_station, by = c('WeatherStation' = 'stid'))

# Preprocessing
pre_ <- preProcess(x = data_solar_description, method = c('center', 'scale'))
data <- predict(object = pre_, newdata = data_solar_description) %>% 
  column_to_rownames('WeatherStation')
```

## Selecting the number of clusters

https://www.datanovia.com/en/lessons/determining-the-optimal-number-of-clusters-3-must-know-methods/

```{r}
# # Elbow method
# p_elbow <- fviz_nbclust(data, kmeans, method = "wss") +
#     geom_vline(xintercept = 4, linetype = 2)+
#   labs(subtitle = "Elbow method")
# 
# # Silhouette method
# p_silhouette <- fviz_nbclust(data, kmeans, method = "silhouette")+
#   labs(subtitle = "Silhouette method")
# 
# # Gap statistic
# set.seed(123)
# p_gap_statistic <- fviz_nbclust(data, kmeans, nstart = 25,  method = "gap_stat", nboot = 500)+
#   labs(subtitle = "Gap statistic method")
# 
# ggarrange(ggarrange(p_elbow, p_silhouette, ncol = 2), p_gap_statistic, nrow = 2)
```

## Plot the clusters of stations
Selected `{r num_clusters}` clustes

```{r}
# Compute k-means clustering with k = num_clusters
set.seed(123)
data_clusters <- kmeans(data, num_clusters, nstart = 25)
fviz_cluster(data_clusters, data = data)

# Cleanup
rm(list = c('data', 'p_elbow', 'p_silhouette', 'p_gap_statistic', 'pre_', 'data_solar_add_train_'))
```

# Train, validation, test and predict split

```{r}
data_solar_add_train_ <- data_solar_add[1:5113, ]

# row indices for training data (70%)
nrow_train <- round(nrow(data_solar_add_train_)*.7, 0)
# row indices for validation data (15%)
nrow_val <- round(nrow(data_solar_add_train_)*.15, 0)
# row indices for test data (15%), the reminder rows
nrow_test <- nrow(data_solar_add_train_)-nrow_train-nrow_val

data_solar_add_train <- data_solar_add_train_[1:nrow_train, ]
data_solar_add_val <- data_solar_add_train_[(nrow_train+1):(nrow_train+nrow_val), ]
data_solar_add_test <- data_solar_add_train_[(nrow_train+nrow_val+1):nrow(data_solar_add_train_), ]

rm(list=c('nrow_train', 'nrow_val', 'nrow_test', 'data_solar_add_train_', 'data_add_col_dates'))
```

# Functions

## Assign the clusters

```{r}
cluster_mean <- function(cluster_num, data = data_solar_add_train) {
  clustered <- names(data_clusters$cluster[data_clusters$cluster == cluster_num])
  
  mean_ <- data %>% 
    select('Date2', all_of(clustered)) %>% 
    pivot_longer(cols = all_of(clustered), names_to = 'WeatherStation', values_to = 'Value') %>%
    group_by(Date2) %>% 
    summarise(Cluster = mean(Value))
  
  data <- mean_ %>% 
   bind_cols(
     data %>%
       select(all_of(c(data_solar_col_predi, data_add_col))))
  
  return(list(
    clustered = clustered,
    mean = mean_,
    data = data
  ))
}
```

## Name to the clustes after an integer

```{r}
cluster_name <- function(x){
  paste('Cluster', x, sep = "_")
}
```

## Add lag to the datasets

```{r}
add_lag <- function(data, num_lag, num_times, col = 'Cluster') {
  col_list <- NULL
  for (i in 1:num_times){
      varname <- paste(col, "lag", num_lag*i , sep="_")
      col_list <- c(col_list, varname)
      data[[varname]] <- lag(data[[col]], num_lag*i)
   }
  data <- data[((num_lag*num_times)+1):nrow(data), ]
  return(list(
    data=data,
    columns=col_list))
}
# add_lag(data_solar_add_train[1:10, 1:3], 2, 3, 'ACME')
```

# Variable importance
Using 'filterVarImp'

```{r}
# cl<-makeCluster(detectCores())
# registerDoParallel(cl)
# 
# select_important<-function(dat, y, n_vars=ncol(dat)){
#   varimp <- filterVarImp(x = dat, y=y, nonpara=TRUE)
#   varimp <- data.table(variable=rownames(varimp),imp=varimp[, 1])
#   varimp <- varimp[order(-imp)]
#   selected <- varimp$variable[1:n_vars]
#   return(selected)
# }
# 
# time_importance <- system.time({
# data_col_importance <- foreach (cluster = 1:num_clusters, .packages=(.packages())) %dopar% {
#                                     data <- cluster_mean(cluster, data_solar_add_train) %>% 
#                                       .$data %>% 
#                                       add_lag(num_lag, num_times, col='Cluster')
# 
#                                     out <- select_important(dat=select(data$data, all_of(c(data_solar_col_predi, data_add_col, data$columns))),
#                                                      y = data$data$Cluster)
#                                     return(out)
#                                     }
# })
# names(data_col_importance) <- cluster_name(1:num_clusters)
# 
# # data_col_importance[2]
# print(time_importance)
# stopCluster(cl)

# saveRDS(data_col_importance, file.path('storage', 'data_col_importance4_lag.rds'))
data_col_importance <- readRDS('~/Downloads/data_col_importance4_lag.rds')
# data_col_importance[1]
# rm(list=c('cl', 'select_important', 'time_importance', 'data'))
```

# Hyperparameter optimization
Using caret library, for all the 'WeatherStations'
Caret
https://topepo.github.io/caret/data-splitting.html
https://topepo.github.io/caret/available-models.html
https://github.com/topepo/caret/blob/master/models/files/svmRadialSigma.R
https://github.com/topepo/caret/blob/master/models/files/xgbLinear.R
https://github.com/topepo/caret/blob/master/models/files/xgbDART.R

Others
https://xgboost.readthedocs.io/en/latest/parameter.html
https://github.com/lucaseustaquio/ams-2013-2014-solar-energy/blob/master/ams-2013-2014-R/
https://stackoverflow.com/questions/30233144/time-series-splitting-data-using-the-timeslice-method
https://robjhyndman.com/hyndsight/tscv/
http://www.quintuitive.com/2016/09/25/better-model-selection-evolving-models/
https://machinelearningmastery.com/pre-process-your-dataset-in-r/

```{r}
model_result <- function(model_train, data_train, data_val) {
  # Get model predictions
  predictions_train <- predict(model_train, newdata = data_train)
  predictions_val <- predict(model_train, newdata = data_val)
  # Get errors
  errors_train <- predictions_train - data_train$Cluster
  errors_val <- predictions_val - data_val$Cluster
  # Compute Metrics
  mse_train <- mean(errors_train^2)
  mae_train <- mean(abs(errors_train))
  mse_val <- mean(errors_val^2)
  mae_val <- mean(abs(errors_val))
  # Personal metrics
  mae_ratio = mae_val/mae_train
  fitting = ifelse(mae_ratio<1, 'Underfitting', ifelse(mae_ratio==0, 'Fit', 'Overfitting'))
  result_combined <- tibble(mse_train, mse_val, mae_train, mae_val, mae_ratio, fitting)
  # Quick plot for 'Test'
  result_plot <- ggplot() +
    geom_point(aes(y=predictions_val, x=1:nrow(data_val)), color = 'red') +
    geom_point(aes(y=data_val$Cluster, x=1:nrow(data_val)), color = 'blue')
  
  return(list(
    mse_train = mse_train, 
    mse_val = mse_val, 
    mae_train = mae_train, 
    mae_val = mae_val, 
    mae_ratio = mae_ratio, 
    fitting = fitting,
    combined = result_combined,
    plot = result_plot
  ))
}

# Define grid seach
cost_values <- 10^seq(from = -2, to = 1, length.out = 5)
epsilon_values <- 10^seq(from = -2, to = 0, length.out = 5)
gamma_values <- 10^seq(from = -3, to = -1, length.out = 5)

# cluster = 1
# cost=epsilon=gamma=1

model_train <- function(cluster, num_col_importance, num_lag, num_times) {
  timestamp()
  
  # Columns to use, depending on the 'num_col_importance'
  col_importance <- data_col_importance[[cluster_name(cluster)]][1:num_col_importance]
  # Subset selection
  data_train <- cluster_mean(cluster, data_solar_add_train) %>% 
    .$data %>% 
    add_lag(num_lag, num_times, col='Cluster') %>%
    .$data %>% 
    select('Cluster', all_of(col_importance))
  data_val <- cluster_mean(cluster, data_solar_add_val) %>% 
    .$data %>% 
    add_lag(num_lag, num_times, col='Cluster') %>%
    .$data %>% 
    select('Cluster', all_of(col_importance))
  
  # Preprocessing
  pre_ <- preProcess(x = select(data_train, -Cluster), method = c('center', 'scale'))
  data_train <- predict(object = pre_, newdata = data_train)
  data_val <- predict(object = pre_, newdata = data_val)

  # Call the training
  trained <- foreach(cost = cost_values, .combine = rbind) %:%
    foreach(epsilon = epsilon_values, .combine = rbind) %:%
      foreach(gamma = gamma_values, .combine = rbind, .packages=(.packages()), .export=ls(envir=globalenv())) %dopar% {
        
        # Model training
        model <- svm(Cluster ~ ., data = data_train, cost = cost, epsilon = epsilon, gamma = gamma)
        # Results
        model_result_ <- model_result(model, data_train, data_val)

        return(data.table(
            cost = cost,
            epsilon = epsilon,
            gamma = gamma,
            mse_train = model_result_$mse_train, 
            mse_val = model_result_$mse_val,
            mae_train = model_result_$mae_train, 
            mae_val = model_result_$mae_val,
            mae_ratio = model_result_$mae_ratio,
            fitting = model_result_$fitting))
  }
  
  return(list(
    num_col_importance = num_col_importance, 
    num_lag = num_lag,
    num_times = num_times,
    cluster = cluster,
    trained = trained
  ))
}

# Parallel grid search
cl<-makeCluster(detectCores())
registerDoParallel(cl)
  model_trained <- lapply(1:num_clusters, function(x) model_train(x, num_col_importance, num_lag, num_times))
stopCluster(cl)
names(model_trained) <- cluster_name(1:num_clusters)

# Filter the best model
model_best <- lapply(1:num_clusters, function(x) {model_trained[[cluster_name(x)]] %>% 
    .$trained %>%
    arrange(desc(mae_val), desc(mae_train)) %>%
    dplyr::slice(1)
})
names(model_best) <- cluster_name(1:num_clusters)

# sapply(1:num_clusters, function(x) model_best[[cluster_name(x)]])

# str(model_trained)
saveRDS(model_trained, file.path('storage', 'model_trained_25_svm.rds'))
# model_trained <- readRDS(file.path('storage', 'model_trained_25_svm.rds'))

rm(model_trained)
# rm(list=c('cl', 'model_train', 'model_result'))
```

# Train and validation final results
Using caret library, for all the 'WeatherStations'

```{r}
model_result <- function(model, data_train_val, data_test) {
  # Get model predictions
  predictions_train_val <- predict(model, newdata = data_train_val)
  predictions_test <- predict(model, newdata = data_test)
  # Get errors
  errors_train_val <- predictions_train_val - data_train_val$Cluster
  errors_test <- predictions_test - data_test$Cluster
  # Compute Metrics
  mse_train_val <- mean(errors_train_val^2)
  mae_train_val <- mean(abs(errors_train_val))
  mse_test <- mean(errors_test^2)
  mae_test <- mean(abs(errors_test))
  # Personal metrics
  mae_ratio = mae_test/mae_train_val
  fitting = ifelse(mae_ratio<1, 'Underfitting', ifelse(mae_ratio==0, 'Fit', 'Overfitting'))
  # Combining the results
  result_combined <- tibble(mse_train_val, mse_test, mae_train_val, mae_test, mae_ratio, fitting)
  # Quick plot for 'Test'
  result_plot <- ggplot() +
    geom_point(aes(y=predictions_test, x=data_test$Date2), color = 'red') +
    geom_point(aes(y=data_test$Cluster, x=data_test$Date2), color = 'blue')
  
  return(list(
    mse_train_val = mse_train_val, 
    mse_test = mse_test, 
    mae_train_val = mae_train_val, 
    mae_test = mae_test, 
    mae_ratio = mae_ratio, 
    fitting = fitting,
    combined = result_combined,
    plot = result_plot
  ))
}

model_apply <- function(cluster, num_col_importance, num_lag, num_times) {
  timestamp()
  
  # Columns to use, depending on the 'num_col_importance'
  col_importance <- data_col_importance[[cluster_name(cluster)]][1:num_col_importance]
  # Subset selection
  data_train <- cluster_mean(cluster, data_solar_add_train) %>% 
    .$data %>% 
    add_lag(num_lag, num_times, col='Cluster') %>%
    .$data %>% 
    select('Cluster', all_of(col_importance))
  
  data_val <- cluster_mean(cluster, data_solar_add_val) %>% 
    .$data %>% 
    add_lag(num_lag, num_times, col='Cluster') %>%
    .$data %>% 
    select('Cluster', all_of(col_importance))
  
  data_train_val <- bind_rows(data_train, data_val)
  
  data_test <- cluster_mean(cluster, data_solar_add_test) %>% 
    .$data %>% 
    add_lag(num_lag, num_times, col='Cluster') %>%
    .$data %>% 
    select('Cluster', all_of(col_importance))
  
  # Preprocessing
  pre_ <- preProcess(x = select(data_train_val, -Cluster), method = c('center', 'scale'))
  data_train_val <- predict(object = pre_, newdata = data_train_val)
  data_test <- predict(object = pre_, newdata = data_test)
  
  # Best tuned hyperparameter
  cost <- model_best[[cluster_name(cluster)]]$cost
  epsilon <- model_best[[cluster_name(cluster)]]$epsilon
  gamma <- model_best[[cluster_name(cluster)]]$gamma
  
  # Model training
  model <- svm(Cluster ~ ., data = data_train_val, cost = cost, epsilon = epsilon, gamma = gamma)
  # Results
  model_result_ <- model_result(model, data_train_val, data_test)

  return(list(
      cost = cost,
      epsilon = epsilon,
      gamma = gamma,
      col_importance = col_importance,
      pre_process = pre_,
      model = model,
      mse_train_val = model_result_$mse_train_val, 
      mse_test = model_result_$mse_test, 
      mae_train_val = model_result_$mae_train_val, 
      mae_test = model_result_$mae_test, 
      mae_ratio = model_result_$mae_ratio,
      fitting = model_result_$fitting))
}

# Parallel apply
cl<-makeCluster(detectCores())
registerDoParallel(cl)

system.time({
  model_applied <- lapply(1:num_clusters, function(x) model_apply(x, num_col_importance, num_lag, num_times))
})
names(model_applied) <- cluster_name(1:num_clusters)

stopCluster(cl)

saveRDS(model_applied, file.path('storage', 'model_applied_25_svm.rds'))
# model_applied <- readRDS(file.path('storage', 'model_applied_25_svm.rds'))

# rm(list=c('cl', 'model_apply', 'model_result'))
```

# Submission File
Submission: https://www.kaggle.com/c/ams-2014-solar-energy-prediction-contest/submit
Leaderboard: https://www.kaggle.com/c/ams-2014-solar-energy-prediction-contest/leaderboard#score
Jesús: https://campus.ie.edu/webapps/discussionboard/do/message?action=list_messages&course_id=_114320331_1&nav=discussion_board_entry&conf_id=_251223_1&forum_id=_112829_1&message_id=_4658342_1

```{r}
# data_solar_add_train_ <- data_solar_add[1:5113, ]
# data_solar_add_predict <- data_solar_add[5114:nrow(data_solar_add), c(data_solar_col_dates, data_solar_col_predi, data_add_col)]

predict_fun <- function(data, row_to_predict, cluster) {

  rows_lag <- num_lag*num_times
  col_importance <- model_applied[[cluster_name(cluster)]]$col_importance
  
  data_predict <- data %>% 
    dplyr::slice((row_to_predict-rows_lag):row_to_predict) %>% 
    add_lag(num_lag=num_lag, num_times=num_times, col='Cluster') %>%
    .$data %>% 
    select(all_of(col_importance))
  
  pre_ <- predict(object = model_applied[[cluster_name(cluster)]]$pre_process, newdata = data_predict)
  value_predicted <- predict(object = model_applied[[cluster_name(cluster)]]$model, newdata = pre_)

  return(value_predicted)
  
}


# cluster_ <- 1
# row_to_predict <- row_ <- 5114
# data <- data$data
# https://stackoverflow.com/questions/8753531/repeat-rows-of-a-data-frame-n-times

system.time({
final <- data_solar_add['Date']
  for (cluster_ in 1:num_clusters) {
    data <- cluster_mean(cluster_, data_solar_add)
    for (row_ in 5114:nrow(data_solar_add)) {
      data$data[row_, 'Cluster'] <- predict_fun(data$data, row_, cluster_)
    }
    out <- map_dfc(data$clustered, ~data$data$Cluster) %>% 
      setNames(data$clustered)
    final <- bind_cols(final, out)
  }
})

out <- final %>% 
  dplyr::select('Date', all_of(data_solar_col_produ)) %>% 
  dplyr::slice(5114:n())

saveRDS(out, file.path('storage', 'out_25_xgb.rds'))
# out <- readRDS(file.path('storage', 'out_25_xgb.rds'))

write.table(x = out, file = file.path('storage', 'out.csv'), sep = ',', dec = '.', row.names = FALSE, quote = FALSE)
# 
# rm(list=c('predict_fun', 'out', 'data'))
```

