---
title: "Programming R Workgroup Project: Machine Learning Model"
author: "Group E"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Load libraries

```{r load_libraries, message=FALSE}
# General porpuse
library(tidyverse)
library(data.table)
library(lubridate)

# Descriptive
library(skimr)

# Visualization
library(ggplot2)

# Machine learning
library(e1071)
library(caret)
# library(cvTools)

# Calculations
library(forecast)
library(mice)
library(outliers) 

# Paralel computing
library(foreach)
library(doParallel)
```

# Load data

```{r load_data}
data_solar <- readRDS(file = file.path('data', 'solar_dataset.RData'))
data_station <- fread(file = file.path('data', 'station_info.csv'))
data_add <- readRDS(file = file.path('data', 'additional_variables.RData'))
```

## Transform data

```{r}
# Source dataset
data_solar <- data_solar[j = Date2 := as.Date(x = Date, format = "%Y%m%d")]

# Add date conversions
data_solar <- data_solar %>% 
  mutate(Year = year(Date2),
         Month = month(Date2, label = TRUE),
         Day = day(Date2),
         Day_Of_Year = yday(Date2),
         Day_Of_Week = wday(Date2, label = TRUE, week_start = 1),
         Days_Since_Origin = time_length(interval(origin, Date2), unit = 'day')) %>% 
  as.data.table(.)

# Columns defined from the enunciate
data_solar_col_produ <- colnames(data_solar)[2:99]
data_solar_col_predi <- colnames(data_solar)[100:456]
data_solar_col_dates <- setdiff(colnames(data_solar), c(data_solar_col_produ, data_solar_col_predi))

# Columns defined from the enunciate
data_add_col <- colnames(data_add)[2:101]
data_add_col_dates <- setdiff(colnames(data_add), data_add_col)
```

> principal_weather_station
 [1] "KENT" "BOIS" "HOOK" "HOLL" "GOOD" "TIPT" "BEAV" "ARNE" "CAMA" "CHEY" "ERIC" "MANG" "WOOD" "BUTL" "BUFF" "BESS" "RETR" "SLAP" "HOBA" "FREE" "RING"
[22] "ALTU" "WEAT" "MEDI" "WAUR" "CHER" "APAC" "WATO" "SEIL" "FTCB" "HINT" "PUTN" "ACME" "MAYR" "LAHO" "KETC" "ELRE" "SULP" "NINN" "BREC" "WASH" "MINC"
[43] "BYAR" "CHIC" "GUTH" "FAIR" "SPEN" "MARE" "BURN" "MADI" "SHAW" "REDR" "DURA" "PERK" "PAUL" "MEDF" "ADAX" "NEWK" "PAWN" "HUGO" "BURB" "STIL" "WYNO"
[64] "FORA" "HASK" "BLAC" "CENT" "BOWL" "BRIS" "OILT" "WILB" "CHAN" "NOWA" "BIXB" "OKEM" "LANE" "TISH" "SKIA" "STUA" "COPA" "IDAB" "TAHL" "MIAM" "VINI"
[85] "SALL" "EUFA" "STIG" "MCAL" "JAYX" "COOK" "PRYO" "WIST" "CLOU" "OKMU" "TALI" "CLAY" "WEST" "MTHE"

## Complete data_add

```{r}
data <- select(data_add, !!data_add_col)

m_ <- 5
maxit_ <- 5
# data_mice_ <- mice(data, m=m_, maxit=maxit_, meth='pmm', seed=500)
# saveRDS(data_mice_, file.path('storage', 'data_add_mice.rds'))
data_mice_ <- readRDS(file.path('storage', 'data_add_mice.rds'))
# summary(data_mice_)

# Average of all the Multivariate Imputation
data_mice <- 0
for (i in 1:m_) data_mice <- data_mice + complete(data_mice_, i)
data_mice <- data_mice/m_

data_add_mice <- bind_cols(select(data_add, !!data_add_col_dates), data_mice)

# Cleanup
rm(list = c('data', 'data_mice_', 'm_', 'maxit_', 'i', 'data_mice', 'data_add'))
```

# Join datasets

```{r}
data_solar_add <- data_solar %>% 
  left_join(data_add_mice, by = 'Date', suffix = c(".solar", ".add"))

rm(list = c('data_solar', 'data_add_mice'))
# skim(data_solar_add)
```

# Train, validation, test and predict split

```{r}
data_solar_add_train_ <- data_solar_add[1:5113, ]
data_solar_add_predict <- data_solar_add[5114:nrow(data_solar_add), c(data_solar_col_dates, data_solar_col_predi, data_add_col)]

# row indices for training data (70%)
nrow_train <- round(nrow(data_solar_add_train_)*.7, 0)
# row indices for validation data (15%)
nrow_val <- round(nrow(data_solar_add_train_)*.15, 0)
# row indices for test data (15%), the reminder rows
nrow_test <- nrow(data_solar_add_train_)-nrow_train-nrow_val

data_solar_add_train <- data_solar_add_train_[1:nrow_train, ]
data_solar_add_val <- data_solar_add_train_[(nrow_train+1):(nrow_train+nrow_val), ]
data_solar_add_test <- data_solar_add_train_[(nrow_train+nrow_val+1):nrow(data_solar_add_train_), ]

rm(list=c('nrow_train', 'nrow_val', 'nrow_test', 'data_solar_add_train_', 'data_solar_add', 'data_add_col_dates'))
```

# Variable importance
Using 'filterVarImp'

```{r}
# cl<-makeCluster(detectCores())
# registerDoParallel(cl)
# 
# select_important<-function(dat, y, n_vars=ncol(dat)){
#   varimp <- filterVarImp(x = dat, y=y, nonpara=TRUE)
#   varimp <- data.table(variable=rownames(varimp),imp=varimp[, 1])
#   varimp <- varimp[order(-imp)]
#   selected <- varimp$variable[1:n_vars]
#   return(selected)
# }
# 
# time_importance <- system.time({
# data_col_importance <- foreach (x = data_solar_col_produ, .errorhandling="remove", .packages=(.packages())) %dopar% {
#                                     select_important(dat=select(data_solar_add_train, !!c(data_solar_col_predi, data_add_col)),
#                                                      y = data_solar_add_train[[x]])
#                                     }
# })
# names(data_col_importance) <- data_solar_col_produ
# 
# print(time_importance)
# stopCluster(cl)

# saveRDS(data_col_importance, file.path('storage', 'data_col_importance1.rds'))
data_col_importance <- readRDS(file.path('storage', 'data_col_importance1.rds'))

rm(list=c('cl', 'select_important', 'time_importance'))
```

# Hyperparameter optimization
Using caret library, for all the 'WeatherStations'
https://topepo.github.io/caret/available-models.html
https://topepo.github.io/caret/data-splitting.html
https://github.com/lucaseustaquio/ams-2013-2014-solar-energy/blob/master/ams-2013-2014-R/
https://github.com/topepo/caret/blob/master/models/files/svmRadialSigma.R
https://stackoverflow.com/questions/30233144/time-series-splitting-data-using-the-timeslice-method
https://robjhyndman.com/hyndsight/tscv/
http://www.quintuitive.com/2016/09/25/better-model-selection-evolving-models/

```{r}
model_result <- function(model_train, data_train, data_val) {
  # Get model predictions
  predictions_train <- predict(model_train, newdata = data_train)
  predictions_val <- predict(model_train, newdata = data_val)
  # Get errors
  errors_train <- predictions_train - data_train$WeatherStation
  errors_val <- predictions_val - data_val$WeatherStation
  # Compute Metrics
  mse_train <- mean(errors_train^2)
  mae_train <- mean(abs(errors_train))
  mse_val <- mean(errors_val^2)
  mae_val <- mean(abs(errors_val))
  # Personal metrics
  mae_ratio = mae_val/mae_train
  fitting = ifelse(mae_ratio<1, 'Underfitting', ifelse(mae_ratio==0, 'Fit', 'Overfitting'))
  # Combining the results
  result_combined <- tibble(mse_train, mse_val, mae_train, mae_val, mae_ratio, fitting)
  # Quick plot for 'Test'
  result_plot <- ggplot() +
    geom_point(aes(y=predictions_val, x=1:nrow(data_val)), color = 'red') +
    geom_point(aes(y=data_val$WeatherStation, x=1:nrow(data_val)), color = 'blue')
  
  return(list(
    combined = result_combined,
    plot = result_plot
  ))
}

# WeatherStation <- 'ACME'
# col_importance_number <- 40

model_train <- function(WeatherStation, col_importance_number = 40) {
  # Columns to use, depending on the 'col_importance_number'
  col_importance <- data_col_importance[[WeatherStation]][1:col_importance_number]
  # Subset selection
  data_train <- bind_cols(WeatherStation = data_solar_add_train[[WeatherStation]],
                    select(data_solar_add_train, !!col_importance))
  data_val <- bind_cols(WeatherStation = data_solar_add_val[[WeatherStation]],
                    select(data_solar_add_val, !!col_importance))

  # Control
  timeControl <- trainControl(method = "timeslice",
                              initialWindow = 365*5,
                              horizon = 365*1,
                              fixedWindow = TRUE,
                              skip = 365*3,
                              allowParallel = TRUE)
  
  # Grid search
  tuneControl <- expand.grid(C = 10^seq(from = -2, to = 1, by = 0.1),
                            sigma = 10^seq(from = -3, to = -1, by = 0.1))

  # Model
  train_time <- system.time({
    model <- caret::train(WeatherStation ~ .,
                      data = data_train,
                      method = "svmRadialSigma",
                      metric = "RMSE",
                      # tuneLength = 20,
                      tuneGrid = tuneControl,
                      trControl = timeControl,
                      preProcess = c('center', 'scale'))

  })

  # Results
  model_result <- model_result(model, data_train, data_val)
  
  # To check the status
  print(paste(WeatherStation, Sys.time(), train_time[3], model_result$combined$mae_ratio, sep = ' || '))

  return(list(
    WeatherStation = WeatherStation,
    col_importance = col_importance,
    train_time = train_time,
    model = model,
    result = model_result
  ))
}

# Parallel grid search
cl<-makeCluster(detectCores())
registerDoParallel(cl)

system.time({
  model_trained <- lapply(data_solar_col_produ, model_train)
})
names(model_trained) <- data_solar_col_produ

stopCluster(cl)

# saveRDS(model_trained, file.path('storage', 'model_trained15.rds'))
# model_trained <- readRDS(file.path('storage', 'model_trained5.rds'))

rm(list=c('cl', 'model_train', 'model_result'))
```

## Results

```{r}
data_solar_col_produ %>% 
  map_dfr(function(x) bind_cols(WeatherStation = x, model_trained[[x]]$result$combined)) %>% 
  ggplot() + 
  geom_histogram(aes(mae_ratio, fill = fitting))
```


# Train and validation final results
Using caret library, for all the 'WeatherStations'

```{r}
model_result <- function(model_train, model_train_val, data_train, data_val, data_train_val, data_test) {
  # Get model predictions
  predictions_train <- predict(model_train, newdata = data_train)
  predictions_val <- predict(model_train, newdata = data_val)
  predictions_train_val <- predict(model_train_val, newdata = data_train_val)
  predictions_test <- predict(model_train_val, newdata = data_test)
  # Get errors
  errors_train <- predictions_train - data_train$WeatherStation
  errors_val <- predictions_val - data_val$WeatherStation
  errors_train_val <- predictions_train_val - data_train_val$WeatherStation
  errors_test <- predictions_test - data_test$WeatherStation
  # Compute Metrics
  mse_train <- mean(errors_train^2)
  mae_train <- mean(abs(errors_train))
  mse_val <- mean(errors_val^2)
  mae_val <- mean(abs(errors_val))
  mse_train_val <- mean(errors_train_val^2)
  mae_train_val <- mean(abs(errors_train_val))
  mse_test <- mean(errors_test^2)
  mae_test <- mean(abs(errors_test))
  # Personal metrics
  mae_ratio = mae_test/mae_train_val
  fitting = ifelse(mae_ratio<1, 'Underfitting', ifelse(mae_ratio==0, 'Fit', 'Overfitting'))
  # Combining the results
  result_combined <- tibble(mse_train, mse_val, mse_train_val, mse_test, mae_train, mae_val, mae_train_val, mae_test, mae_ratio, fitting)
  # Quick plot for 'Test'
  result_plot <- ggplot() +
    geom_point(aes(y=predictions_test, x=data_test$Date2), color = 'red') +
    geom_point(aes(y=data_test$WeatherStation, x=data_test$Date2), color = 'blue')
  
  return(list(
    combined = result_combined,
    plot = result_plot
  ))
}

model_apply <- function(WeatherStation, col_importance_number = 40) {
  
  # Final model trained
  trained_ <- model_trained[[WeatherStation]]
  
  # Final coefficients
  model_coef <- data.table(
    # epsilon = trained_$model$finalModel@param$epsilon,
    C = trained_$model$finalModel@param$C,
    sigma = trained_$model$finalModel@kernelf@kpar$sigma
  )
  
  # Columns to use, depending on the 'col_importance_number'
  col_importance <- trained_$col_importance
  
  # Subset selection
  data_train <- bind_cols(WeatherStation = data_solar_add_train[[WeatherStation]],
                    select(data_solar_add_train, !!col_importance))
  data_val <- bind_cols(WeatherStation = data_solar_add_val[[WeatherStation]],
                    select(data_solar_add_val, !!col_importance))
  data_train_val <- bind_rows(data_train, data_val)
  data_test <- bind_cols(WeatherStation = data_solar_add_test[[WeatherStation]], 
                    select(data_solar_add_test, 'Date', 'Date2', !!col_importance))
  
  # Control
  timeControl <- trainControl(method = "timeslice",
                              initialWindow = 365*5,
                              horizon = 365*1,
                              fixedWindow = TRUE,
                              skip = 365*3,
                              allowParallel = TRUE)

  # Model
  train_time <- system.time({
    model <- caret::train(WeatherStation ~ .,
                      data = data_train,
                      method = "svmRadialSigma",
                      metric = "RMSE",
                      tuneGrid = model_coef,
                      trControl = timeControl,
                      preProcess = c('center', 'scale'))
  })
  
  # Results
  model_result <- model_result(trained_$model, model, data_train, data_val, data_train_val, data_test)

    # To check the status
  print(paste(WeatherStation, Sys.time(), train_time[3], model_result$combined$mae_ratio, sep = ' || '))

  return(list(
    WeatherStation = WeatherStation,
    train_time = train_time,
    model = model,
    result = model_result
  ))
}

# Parallel apply
cl<-makeCluster(detectCores())
registerDoParallel(cl)

system.time({
  model_applied <- lapply(data_solar_col_produ, model_apply)
})
names(model_applied) <- data_solar_col_produ

stopCluster(cl)

saveRDS(model_applied, file.path('storage', 'model_applied15.rds'))
# model_applied <- readRDS(file.path('storage', 'model_applied15.rds'))

rm(list=c('cl', 'model_apply', 'model_result'))
```

## Results

```{r}
data_solar_col_produ %>% 
  map_dfr(function(x) bind_cols(WeatherStation = x, model_applied[[x]]$result$combined)) %>% 
  ggplot() + 
  geom_histogram(aes(mae_ratio, fill = fitting))
```

# Submission File
Submission: https://www.kaggle.com/c/ams-2014-solar-energy-prediction-contest/submit
Leaderboard: https://www.kaggle.com/c/ams-2014-solar-energy-prediction-contest/leaderboard#score
Jesús: https://campus.ie.edu/webapps/discussionboard/do/message?action=list_messages&course_id=_114320331_1&nav=discussion_board_entry&conf_id=_251223_1&forum_id=_112829_1&message_id=_4658342_1

```{r}
system.time({
  predict_ <- data_solar_col_produ %>% 
    setNames(nm = .) %>% 
    map_dfc(function(WeatherStation) predict(object = model_applied[[WeatherStation]]$model, newdata = data_solar_add_predict))
})

out <- tibble(Date = data_solar_add_predict$Date, predict_)
write.table(x = out, file = file.path('storage', 'out.csv'), sep = ',', dec = '.', row.names = FALSE, quote = FALSE)

rm(list=c('predict_', 'out'))
```

