---
title: "Solar Production EDA"
author: "Group E"
date: "4/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

This report contains the findings on Solar data, additional data and stations data. 

# Preparation

###### 1 Load libraries

```{r load_libraries, message=FALSE}
# General purpose
library(tidyverse)
library(data.table)
library(lubridate)

# Descriptive
library(skimr)

# Visualization
library(ggplot2)
library(PerformanceAnalytics)
library(corrplot)

# Mapping
library(leaflet)
library(leaflet.extras)
library(sf)
#library(gdtools)

# Calculations
library(forecast)
library(caret)
library(mice)
library(outliers) 

library(foreach)
library(doParallel)
```

###### 2 Source From R script
```{r}

source('20200411_Main_EDA.R', echo = FALSE, print.eval = FALSE)

```

###### 3 Read prepared Datasets
```{r,eval=FALSE}
# Solar production dataset 
data_solar_train
data_solar_test
data_station 
data_add

```

# Overview: Data structure and content

##### 1 General overview

##### 1.1 NA check

```{r}
f_check_na(data_solar_train)
f_check_na(data_solar_test)
f_check_na(data_station)
f_check_na(data_add)

```
> Comment: The missing values in data_add will be imputed later

##### 1.2 Constant Variables check
```{r}
criteria_variables(data_solar_train)
criteria_variables(data_solar_test)
criteria_variables(data_station)
criteria_variables(data_add)
```
> Comment: No constant variables have been found.

##### 1.3 Outliers check

Based on:
https://www.kaggle.com/rtatman/data-cleaning-challenge-outliers
https://cran.r-project.org/web/packages/outliers/outliers.pdf

###### 1.3.1 Outliers in Solar dataset

```{r, eval=FALSE}
library(knitr)
# solar_data_outlier_skim
plot(solar_data_outlier_hist)

kable(solar_data_outlier_table)

```
> The data is normally distributed. There is only one row that could be considered outliers. 

###### 1.3.2Outliers in Additional bariables
First, need to complete the data set to be able to use 'outliers()'.

####  2 Exploring the columns of Solar dataset 

##### 2.1 Descriptive statistics of the Training set

```{r, eval=FALSE}
skim(data_solar_train)
glimpse(data_solar_train)
```

#####2.2Descriptive statistics of the Testing set

```{r, eval=FALSE}
skim(data_solar_test)
glimpse(data_solar_test)
```

#####2.3 Descriptive statistics of the Stations dataset

```{r, eval=FALSE}
skim(data_station)
glimpse(data_station)
```

### Descriptive statistics of Additional information dataset

```{r, eval=FALSE}
skim(data_add)
glimpse(data_add)
```

# Descriptive plots

## Principal weather stations
Firstly, weather stations are sorted by their respective solar energy production volume
```{r}

principal_weather_station

```
Then, we create 2 plots, the first one shows top 7 weather stations

?? What conclusion can we draw
```{r}
multiplot(p_top, p_bottom)
```
## Rank position change over time
We are interested to find out how does the production changes over time. By looking at the top 10 stations. 

As demonstrated by the graph below, there are quite a lot of movements for every weather station. But for one weather station, they typically won't move more than 5 ranking. 

```{r}
p_rank
```
## Total production seasonality, trend and cyclical movements 

Descriptive plot to show the production recorded in millions over different granularities of time.

```{r}
multiplot(p_all, p_year, p_month, p_Day_Of_Week, layout=layout)
```

## Evolution of principal weather stations
Plot for the top highest weather stations.
```{r}
p_top_five_ts
```
## What about Training and Test?
```{r}
p_train_test
```

## Seasonality
https://towardsdatascience.com/forecasting-with-r-trends-and-seasonality-def24280e71f
```{r}
p_seasonality_all
```

### Seanonality of the top 5 stations

```{r}
p_seasonality_top_5
```

# Geolocalization

## Positions - Where are the stations located?
```{r}
map_stations
```

## Heatmap - visualizes production volume
```{r}
map_production
```

# Correlations 
## What can we see from the top 5 stations with the top 5 principal components?
```{r}
p_corr
```

# Histograms, density kernel and boxplot

## Of general production
```{r}
p_distri_prod
```

## Of each weather station 
```{r}
p_histogram_density

p_boxplot
```
## Of predictors 
```{r}
predictor_density
predictor_boxplot
```

# Variable Importance 

Computed Variable Importance is stored in data_solar_importance
```{r}
names(data_solar_importance)
```
> Note: used parallel computing to reduce run time 

# Additional information

## Feature visualization
```{r}
p_additional
```
## Correlation in additional dataset
```{r}
p_corrplot
```
## Impute missing values in additional dataset
```{r}
complete_mice
```
## Outliers in additional dataset
```{r}
add_outlier
add_outlier_hist
add_outlier_table 
```

## PCA of additional dataset

```{r}
data_add_col_pca
data_add_pca
```
### Plot density and boxplots
```{r}
pca_histogram_density
pca_boxplot

```
