---
title: "Solar Production EDA"
author: "Group E"
date: "4/25/2020"
output: html_document
---

```{r setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(message = FALSE)
```

This report contains the findings on Solar data, additional data and stations data. 

## Preparation

###### 1 Load libraries

```{r load_libraries, message=FALSE}
# General purpose
library(tidyverse)
library(data.table)
library(lubridate)

# Descriptive
library(skimr)

# Visualization
library(ggplot2)
library(PerformanceAnalytics)
library(corrplot)

# Mapping
library(leaflet)
library(leaflet.extras)
library(sf)
#library(gdtools)

# Calculations
library(forecast)
library(caret)
library(mice)
library(outliers) 

library(foreach)
library(doParallel)
library(parallel)
```

###### 2 Source From R script
```{r, eval=TRUE, results='hide', include=FALSE}

source('20200411_Main_EDA.R', echo = FALSE, print.eval = FALSE)

```

###### 3 Read prepared Datasets
```{r,eval=FALSE}
# Solar production dataset 
data_solar_train
data_solar_test
data_station 
data_add

```

## Overview: Data structure and content

##### 1 General overview

##### 1.1 NA check

```{r, eval=FALSE}
f_check_na(data_solar_train)
f_check_na(data_solar_test)
f_check_na(data_station)
f_check_na(data_add)

```
> Comment: There are no missing values in the solar_dataset and station dataset. There are 3.6% NAs in the additional_dataset. The missing values will be imputed.

##### 1.2 Constant Variables check
```{r, eval=FALSE}
criteria_variables(data_solar_train)
criteria_variables(data_solar_test)
criteria_variables(data_station)
criteria_variables(data_add)
```
> Comment: No constant variables have been found.

##### 1.3 Outliers check

Based on:
https://www.kaggle.com/rtatman/data-cleaning-challenge-outliers
https://cran.r-project.org/web/packages/outliers/outliers.pdf

##### 1.3.1 Outliers in Solar dataset

```{r}
library(knitr)
# solar_data_outlier_skim
plot(solar_data_outlier_hist)

#kable(solar_data_outlier_table)

```
> The data is normally distributed. There is only one row that could be considered outliers. 

####  2 Exploring the columns of Solar dataset 

##### 2.1 Descriptive statistics of the Training set

```{r, eval=FALSE}
skim(data_solar_train)
glimpse(data_solar_train)
```

##### 2.2 Descriptive statistics of the Testing set

```{r, eval=FALSE}
skim(data_solar_test)
glimpse(data_solar_test)
```

##### 2.3 Descriptive statistics of the Stations dataset

```{r, eval=FALSE}
skim(data_station)
glimpse(data_station)
```

##### 2.4 Descriptive statistics of Additional information dataset

```{r, eval=FALSE}
skim(data_add)
glimpse(data_add)
```

# Descriptive plots

#### 1 Principal weather stations

* In this step, weather stations are sorted by their respective solar energy production volume, in descending order. Then, we create 2 plots, the first one shows top weather stations, the second shows the bottom weather stations. 

* As shown in the graph below, the differences between stations are small.

```{r, eval=TRUE, echo=FALSE}
multiplot(p_top, p_bottom)
```
#### 2 Rank position change over time

* We are interested to find out how does the production of one station changes over years

* As demonstrated by the graph below, there are quite a lot of movements for every weather station. But for each one weather station, they typically won't move more than 5 ranking positions.
+ This suggests variables apart from seasonality at play, which would move every station in the same direction.

```{r, echo=FALSE}
p_rank
```
#### 3 Total production seasonality, trend and cyclical movements 

Descriptive plot to show the production recorded in millions over different granularities of time.

* From 1994 to 2007, the total production volume is slightly trending down.
* There is obvious seasonality factor during the course of one year, i.e. during summer months, the energy production peaks.

```{r, echo=FALSE, message=FALSE}
multiplot(p_all, p_year, p_month, p_Day_Of_Week, layout=layout)
```

#### 4 Dive into Seasonality

This section is based on https://towardsdatascience.com/forecasting-with-r-trends-and-seasonality-def24280e71f

* Decomposing the seasonality factor
* First of all, echoing what we have seen in the above section, there is clear seasonality during a year
* Secondly, the trend over the years is more random, there isn't a pattern (e.g. continuous downward or upward trend) we can observe. Mostly like due to environmental conditions. 

```{r, echo=FALSE}
plot(decompose(data2))
```

#### 5 Visualizing the production timeseries of the top 5 stations

* Not suprisingly, all stations production trends follow the trend isolated in the graph above. 

```{r, echo=FALSE}
top_ <- 5

ts_ <- function(x) {
  a <- ts(data = x, frequency = 365, start = c(1994, 1, 01), end = c(2007, 12, 31))
  a <- decompose(a)
  return(a)
}

data <- data_solar_train %>%
  pivot_longer(cols = all_of(data_solar_col_produ), names_to = 'WeatherStation', values_to = 'Value') %>%
  group_by(WeatherStation) %>%
  group_map(~ ts_(.x$Value)) %>%
  setNames(data_solar_col_produ)

# bb <- lapply(principal_weather_station[1:top_], function(x) plot(data[[x]]))
p_seasonality_top_5 <- lapply(principal_weather_station[1:top_], function(x) plot(data[[x]]$trend, main = x, ylab = 'Value'))
```

## Geolocalization

#### 1 Positions - Where are the stations located?

```{r, echo=FALSE}
map_stations
```

#### 2Heatmap - visualizes production volume
```{r}
map_production
```

## Correlations 
#### 1 Top 5 stations vs first 10 principal components

```{r, echo=FALSE}
top_ <- 5
top_pc <- 10

data <- data_solar_train %>%
  dplyr::select(principal_weather_station[1:top_], data_solar_col_predi[1:top_pc])

chart.Correlation(data, histogram=TRUE) #, pch=19
```

## Histograms, density kernel and boxplot

#### 1 Of general production
```{r, echo=FALSE, message=FALSE}
multiplot(p_histogram_density, p_boxplot, layout = layout)
```

#### 2 Of each weather station 
```{r, echo=FALSE}
data <- data_solar_train %>%
  dplyr::select(data_solar_col_produ) %>%
  pivot_longer(cols = all_of(data_solar_col_produ), names_to = 'WeatherStation', values_to = 'Value') %>%
  mutate(WeatherStation_f = factor(WeatherStation, levels=principal_weather_station))

ggplot(data = data, aes(x = Value/1e6)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="blue") +
  facet_wrap(vars(WeatherStation_f))

ggplot(data = data, aes(x = Value/1e6)) +
  geom_boxplot() +
  stat_boxplot(coef = 1.5, outlier.colour = 'red', outlier.alpha = 0.1) +
  facet_wrap(vars(WeatherStation_f)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```
#### 3 Of predictors 

* Predictors are highly concentrated, the variances are very small
```{r}
data <- data_solar_train %>%
  dplyr::select(data_solar_col_predi) %>%
  pivot_longer(cols = all_of(data_solar_col_predi), names_to = 'PC', values_to = 'Value') %>%
  filter(PC %in% data_solar_col_predi[1:50]) %>%
  mutate(PC_f = factor(PC, levels=data_solar_col_predi))

ggplot(data = data, aes(x = Value)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="blue") +
  facet_wrap(vars(PC_f))

ggplot(data = data, aes(x = Value)) +
  geom_boxplot() +
  stat_boxplot(coef = 1.5, outlier.colour = 'red', outlier.alpha = 0.1) +
  facet_wrap(vars(PC_f)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

# Variable Importance 

> Computed Variable Importance is stored in data_solar_importance. Used parallel computing to reduce run time.

* Below is an example of variable importance ranked descendingly, for the top 5 weather station.  

```{r, echo=FALSE}
top_ <- 5

data_solar_importance <- readRDS(file.path('storage', 'data_solar_importance_parallel.rds'))

names(data_solar_importance) <- principal_weather_station[1:top_]

data_solar_importance[1:top_]

```


## Additional information dataset

#### 1 Distribution of the descriptive stats

```{r, echo=FALSE}
multiplot(p_mean, p_median, p_sd, p_na, p_mean_sd, layout = layout)
```
#### 2 Correlation in additional dataset

```{r, echo=FALSE}
corrplot(cor_, type="upper", order="hclust", p.mat = p.mat, sig.level = 0.01, insig = "blank")
```

#### 3 PCA of additional dataset

```{r, echo=FALSE}
bind_cols(data_add[, ..data_add_col_dates], df4)
```
### Plot density and boxplots
```{r}
data <- data_add_pca %>%
  pivot_longer(cols = all_of(data_add_col_pca), names_to = 'PC', values_to = 'Value') %>%
  mutate(PC_ = factor(PC, levels = data_add_col_pca))

ggplot(data = data, aes(x = Value)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="blue") +
  facet_wrap(vars(PC_))

ggplot(data = data, aes(x = Value)) +
  geom_boxplot() +
  stat_boxplot(coef = 1.5, outlier.colour = 'red', outlier.alpha = 0.1) +
  facet_wrap(vars(PC_)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


```
